Learning a reward function from demonstrations: a cascaded supervised learning approach

### 2012-10-18 piot2012cascading/ECML13_cascading.tex
WIP sur le papier ECML

% Created 2012-10-15 lun. 14:49
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage[integrals]{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\tolerance=1000
\usepackage{amsmath}
\usepackage{dsfont}
\providecommand{\alert}[1]{\textbf{#1}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\title{Outline}
\author{Edouard Klein}
\date{\today}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.11}}

\begin{document}

@@

\setcounter{tocdepth}{3}
\tableofcontents
\vspace*{1cm}
\begin{abstract}
  This paper considers the Inverse Reinforcement Learning (IRL) problem which consists in inferring a reward function for which a demonstrated expert policy is optimal.
We propose to break the IRL problem down into two Supervised Learning (SL) steps: this is the cascaded approach. The first step consists in a classification and the second step in a regression.
This cascaded approach is justified by an analysis which shows that the demonstrated expert policy is near-optimal for the reward function outputted by the regression step. Besides, we provide an algorithm which is an instantiation of the cascaded approach. This algorithm does not need to solve repeatedly the Reinforcement Learning (RL) problem and use only sampled transitions as an input.
Finally, up to the use some heuristics, the algorithm may work with only transitions sampled according to the demonstrated expert policy. This is illustrated on a car driving simulator, on a Gridworld and on an inverted pendulum.
  \end{abstract}
\section{Introduction}
\label{sec-2}

  De la même manière que pour le papier de SCIRL, l'intro est une vaste annonce du papier, avec des références pointues (sous-section, voire sous,sous-section) vers les parties du papier qui expliquent tel ou tel point en détail.
\section{Background and notations}
\label{sec:background}
%% \begin{itemize}
%% \item Parler du domaine de la prise de décision séquentielle
%% \item Expliquer le problème direct, en profiter pour introduire l'équation d'évaluation de Bellman et l'équation d'optimalité de Bellman
%% \begin{itemize}
%% \item $RL(MDP=\{S,A,\gamma,(P),R\}) = \pi^*:S\times A \rightarrow \mathbb{R}$
%% \item $\forall \pi, Q^\pi(s,a) = R(s,a) + \sum_{s'}P(s'|s,a)Q^\pi(s',\pi(s'))$
%% \item $\pi^* = \arg\max_aQ^*(s,a)$
%% \end{itemize}
%% \item Introduire le problème inverse dans une forme mathématique relativement propre
%% \begin{itemize}
%% \item $IRL( MDP\backslash R = \{S,A,\gamma,(P)\} \cup \{\pi^E\} = R : S\times A \rightarrow \mathbb{R}$ telle que $\pi^E$ est optimale pour $R$
%% \item Mentionner la nature mal posée du problème sans s'appesentir dessus. Préciser toutefois qu'aucune approche ne le résoud de manière mathématiquement satisfaisante à cause de ça.
%% \item insister sur l'optimalité de l'expert comme hypothèse. Un expert sous optimal, c'est un autre champ de littérature, éventuellement donner deux trois refs.
%% \end{itemize}
%% \end{itemize}
First, we introduce some general notations.
Let $E$ and $F$ be two non-empty sets, $E^F$ is the set of functions from $F$ to $E$.
Let $X$ be a non-empty finite set. We note $\Delta_X$ the set of distributions over $X$ and $Card(X)$ its cardinal. 
Let $\alpha\in\mathbb{R}^X$ and $\beta\in\mathbb{R}^X$: $\alpha\leq\beta$ is a componentwise inequality.

We frame the sequential decision problem in a finite \emph{Markov Decision Process} (MDP), that is a tuple $\{S,A,P,R,\gamma\}$. $S=\{s_i\}_{1\leq i\leq Card(S)}$ is a finite state space, $A=\{a_i\}_{1\leq i\leq Card(K)}$ is a finite action space, $R\in\mathbb{R}^{S\times A}$ is a reward-function, $\gamma\in ]0,1[$ is a discount factor and $P\in \Delta_{S}^{S\times A}$ is the dynamic of the MDP. Thus for each $(s,a)\in S\times A$, $P(.|s,a)$ is a distribution over $S$ and $P(s'|s,a)=p(s,a,s')$ is the probability to reach $s'$ by choosing the action $a$ in the state $s$. At each time step $t$, the agent uses the information encoded in the state $s_t\in S$ in order to choose an action $a_t \in A$ according to a (deterministic\footnote{We restrict ourselves here to deterministic policies, but this dos not induce any loss of generality.}) \emph{policy} $\pi\in A^S$. The agent then transits to a new state $s_{t+1}\in S$ according to the markovian transition probabilities $P(s_{t+1}|s_t,a_t)$. The reward function $R$ is a local measure of the quality of the control. The global quality of the control induced by a policy $\pi$, with respect to a reward $R$ is assessed by the value function $V^\pi_R \in \mathbb{R}^{S}$ which associates each state with the discounted cumulative reward for following policy $\pi$ from this state: $V^\pi_R(s) = \E[\sum_{t\geq 0}\gamma^tR(s_t)|s_0 = s]$. Therefore, an optimal policy $\pi^*_R$ is a policy whose value function (the optimal value function $V^*_R$) is greater than that of any other policy, for all states : $\forall \pi, \forall s \in S, V^*_R(s) \geq V^\pi_R(s)$. The stationary distribution induced by a policy $\pi$ is written $\rho_\pi,\textrm{ s.t. }\rho_\pi^TP^\pi = \rho_\pi$.

The Bellman evaluation operator $T^\pi_R: \mathbb{R}^{S} \rightarrow  \mathbb{R}^{S}$ is defined by $T^{\pi_R}V = R + \gamma P_\pi V$ using the straightforward matricial notation $P_\pi = (p(s'|s,\pi(s)))_{s,s' \in S}$. The Bellman optimality operator follows naturally : $T^*_RV = \max_\pi T^\pi_RV$. Both operators are contractions. The fixed point of the Bellman evaluation operator $T^\pi_R$ is the value function of $\pi$ with respect to reward R: $V^\pi_R$; the fixed point of the Bellman optimality operator $T^*_R$ is the optimal value function $V_R^*$ with respect to reward $R$. Optimal value function and optimal policy are also linked via the quality function $Q^\pi_R\in\mathbb{R}^{S\times A}$, a state-action value function that adds a degree of freedom on the choice of the first action, formally defined by $Q^\pi_R(s,a) = T^a_RV^\pi_R(s)$, with $a$ the policy that always returns action $a$ ($T^aV = R + \gamma P_a V$ with $P_a = (p(s'|s,a))_{s,s' \in S}$). The optimal policy follows a greedy mechanism with respect to its optimal quality function $\pi^*_R(s)\subset\argmax_aQ^*_R(s,a)$.

When the state space is too large to allow matricial representation or when the transition probability or even the reward function are unknown except through observations gained by interacting with the system, \emph{Reinforcement Learning} (RL) or approximate dynamic programming may be used to retrieve the optimal control policy. We call this the direct problem.

In this work, we wish to tackle the inverse problem. We observe an expert's deterministic\footnotemark[\value{footnote}] policy $\pi^E$, assuming that there exist some unknown reward $R^E$ for which the expert is optimal. The non optimality of the expert is an interesting setting that has been discussed for example FIXME:NONOPTIMALITY but that we are not addressing here. We do not try to find this unknown reward $R^E$ but rather a non trivial reward $R$ for which the expert is also optimal. The trivial reward $0$ is a solution to this ill-posed problem (no reward means that every behavior is optimal). Because of its ill-posed nature, this expression of \emph{Inverse Reinforcement Learning} (IRL) still has to find a satisfactory solution although lots of progress have been made, see section \ref{sec:related}.
\section{The cascading algorithm}
\label{sec:algo}
%% \begin{itemize}
%% \item Mentionner (sans en parler, ce sera fait dans \hyperref[sec-7-1]{Imitation}) l'imitation. Une approche supervisée évidente est la classification, avec pour entrée un set $D_C$ de données issues de l'expert. Rappeler (expliqué pour la première fois dans \hyperref[sec-3]{Formalisme} qu'on se restreint (sans perte de généralité, le préciser) au cas des politique déterministes.
%% \item Tous les classifieurs utilisent une fonction de score (je n'ai pas trouvé un seul exemple d'un classifieur qui n'en utilise pas), il s'agit juste d'un manière d'écrire les choses.
%% \item Expliquer que l'on identifie le mécanisme de classification au mécanisme glouton qui régit la politique optimale (donc celle de l'expert).
%% \item Sans chercher à justifier mathématiquement la chose (ce sera fait dans la partie \hyperref[sec-5]{Analyse}), expliquer qu'intuitivement l'on va identifier fonction de score du classifieur et fonction de qualité de l'expert (qui est optimale pour une certaine récompense, rappelons-le).
%% \begin{itemize}
%% \item Repartir de l'équation d'évaluation de Bellman pour isoler le terme $R(s,a)$ et introduire $r:S\times A\in \mathbb{R}$, la récompense issue de cette équation lorsque l'on remplace $Q^E$ par la fonction de score : $r(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi(s')) $
%% \item On estime ce terme à partir des transitions $D_R$ car les probabilités de transitions sont inconnues : $\hat r(s,a) = q(s,a)-\gamma \sum_{s'} {1\over N_{s'}}\sum_{(s,a,s',a')\in D_R}q(s',a')$
%% \item La fonction de récompense est donnée par une régression sur $D_R \cup {\hat r}$.
%% \end{itemize}
%% \item Donner l'algorithme l'environnement \texttt{algorithm2e}
%% \item Rapidement, l'étape de régression permet d'ajouter la structure du MDP au résultat de classif, en utilisant l'équation de Bellman.
%% \item Bien préciser qu'il ne s'agit ici que d'une explication qualitative, ``avec les mains'', la preuve du fait que ça résoud bien le problème est donnée dans \hyperref[sec-5]{Analyse}, une démonstration sur des problèmes si complexes que l'état de l'art ne peut les résoudre (sauf SCIRL) est donnée dans \hyperref[Experience]{Experience}. Les avantages comparés sont expliqués dans \hyperref[Autres-Approches]{Autres Approches}. Enfin, des explications plus intuitives sur les mécanismes internes de l'algorithme, ainsi que des résultats expérimentaux supplémentaires sont fournis en \hyperref[sec-9]{Annexes}.
%% \end{itemize}
IRL has sometimes FXME:REFIMITATION been seen as a way to cast the imitation problem, that is a way to copy the expert policy ; the reward function is then a byproduct of the analysis and not the object one is looking for. An efficient supervised way to do pure imitation is to use a classifier. The entries are the states $s_t$ and the labels are the actions $a_t$ deterministically chosen by the expert. We thus obtain a classification policy $\pi_C$.

This classification operation is the first step of our algorithm, which consists in cascading two supervised approaches. Using a dataset $D_C = {s_i,a_i=\pi^E(s_i)}_i$ we train a score-function based classifier. We can make use of any classifier whose decision rule uses a greedy mechanism with respect to a score function $q : S \times A \rightarrow \mathbb{R}$ that quantifies how well label $a\in A$ suits entry $s\in S$. Most classifiers, from M-SVM to $k$-nearest neighboors can be cast as a score-function based classifier. This allows the use of many off-the-shelf imlpementations.

The greedy classification mechanism $\pi_C = \argmax_a q(s,a)$ is very similar to the relation between the optimal policy in a MDP and its optimal quality function : $\pi^*_R(s) = \argmax_aQ^*_R(s,a)$. As the expert is optimal with respect to some unknown reward function $R_E$ and we seek a reward function $R$ for which it is also optimal, it is only natural to view the score function $q$ of the classifier as some kind of optimal quality function. It is not a stretch to see $\pi_C$ as an optimal policy as it has been built from $\pi^E$ which is optimal. By inversing the Bellman evaluation equation, one can express the reward as a function of the quality function : $R(s,a) = Q^*_R(s,a) - \sum_{s'}P(s'|s,a)Q^*_R(s'\pi^*(s'))$. This leave us with $r$, the reward function relative to our score/quality function $q$ : $r(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi_C(s'))$.

We wish to solve the general, or approximate IRL problem, where the transition probabilities are unknown. Our reward function $r$ must be approximated to $\hat r$ with the help of a dataset $D_R = \{s_i,a_i,s'_i\}$. This gives as many datapoints $\{\hat r(s_i,a_i) = q(s_i,a_i) - \gamma q(s',\pi_C(s'))\}_i$ that we then feed to a regressor to obtain $R^C$, a generalisation of $\hat r$ over the whole state-action space. The complete algorithm is given Alg. \ref{algo:cascading}.
\begin{algorithm}%[H]
    %\small
  \caption{Cascading IRL algorithm}
  \label{algo:cascading}
  \emph{\textbf{Given}} a training set $D_C=\{(s_i,a_i=\pi^E(s_i))\}_{1\leq i \leq D}$ and another training set $D_R=\{(s_{i},a_{i},s'_{i})\}_{1\leq i \leq D'}$\;\\
  \emph{\textbf{Train}} a score function-based classifier on $D_C$, obtaining decision rule $\pi_C$ and score function $q:S\times A \rightarrow \mathbb R$\;\\
  \emph{\textbf{Learn}} a reward function $\hat R^C$ from the dataset $\{(s_{i},a_{i},\hat{r}_i)\}_{1\leq i \leq D'}$, $\forall (s_i,a_i,s'_i) \in D_R,\hat{r}_i=q(s_{i},a_{i})-\gamma q(s'_{i},\pi_C(s'_{i}))$\;\\
  \emph{\textbf{Output}} the reward function $\hat R^{C}$ \;
\end{algorithm}

Cascading two supervised approaches like we do is a way to inject the MDP structure in the resolution of the problem. Indeed, mere classification only takes into account information from the expert (i.e., which actions goes with what state), but there is more to a MDP than a single policy. Using the Bellman equation in the expression of $\hat r$, we make use of the information lying in the transitions $(s,a,s')$, namely information about the transition probability $P$. The final regression step is just a way to try to generalize this information about $P$ to the whole state-action space, in order to have a well-behaved reward function. As we make direct use of the available data and do not iteratively compare some metric of the stationary distribution of policies, we do not need to solve the forward problem. This is a huge advantage, as it greatly reduce the computational and sampling cost of the method.

By carefully choosing or tuning both supervised algorithm, it should possible to mitigate ill effects of a lack of data. The classification error and regression error can be acted upon with the help of the wide range of techniques available for these two well-known supervised problems, which helps control the bound we provide in Sec. \ref{sec:analysis}. The explanation in this section only is a quatitative attempt to make the reader understand the idea behind the algorithm. The analysis tackle things in a formal, yet less intuitive, way. Dire lack of data is addressed in Sec \ref{sec:experiments} where the use of an heuristic is empirically shown to be relevant. Data constraints are so high that to the best of our knowledge only one other published algorithm is able to solve the IRL problem in such conditions. Other approaches are compared with one another in Sec. \ref{sec:related}. Finally, appendices provide the reader with other experiments that exemplify the inner working of the algorithm.

@@

First let introduce some notations. $\pi_C$ is the classifier policy, $\epsilon_C$ is the classification error:

@@

@@

@@

Now, there is a still an upper bound to find for the second term. If we introduce, $R^E_a=q_a+\gamma P_aq_{\pi_E}$, it is possible to decompose the second term:

>>>>

% Created 2012-10-15 lun. 14:49
\documentclass[smallextended]{svjour3}
\smartqed 
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage[integrals]{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\tolerance=1000
\usepackage{amsmath}
%\usepackage{amsthm}
%\newtheorem{definition}{Definition}
%\newtheorem{theorem}{Theorem}
\usepackage{dsfont}
\providecommand{\alert}[1]{\textbf{#1}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\begin{document}
\title{Learning a reward function from demonstrations: a cascaded supervised learning approach}
\author{Edouard Klein$^{1,2}$ \and Bilal Piot $^{2,3}$\and Matthieu Geist $^{2}$\and Olivier Pietquin$^{2,3}$}
\institute{
 1. ABC Team\\
 LORIA-CNRS, France.\\
\and
2.Supélec-Metz Campus\\
 MaLIS Research group, France\\
 \texttt{firstname.lastname@supelec.fr}\\
\and 
3. UMI 2958 CNRS\\
GeorgiaTech, France\\
}
\date{\today}

@@

\begin{abstract}
  This paper considers the Inverse Reinforcement Learning (IRL) problem, that is inferring a reward function for which a demonstrated expert policy is optimal.
We propose to break the IRL problem down into two Supervised Learning (SL) steps: this is the cascaded approach. A classification step aiming to imitate the expert's policy gives a score function we see as an optimal action value function. In the second step, this score function is used together with the bellman evaluation equation to create samples then fed to a regressor that outputs a reward function.
This cascaded approach is justified by an analysis which shows that the demonstrated expert policy is near-optimal for the reward function the regression step outputs. We explain some advantages of this approach over existing algorithms, mainly not needing to repeatedly solve the Reinforcement Learning (RL) direct problem.
Finally, up to the use some heuristics, we empirically show the algorithm to work with only transitions sampled according to the demonstrated expert policy.
  \end{abstract}
\section{Introduction}
\label{sec-2}

Inverse Reinforcement Learning (IRL)~\cite{russell1998learning} aims at inferring a reward function for which a demonstrated expert policy is optimal. An important part of the literature (see Section~\ref{sec:related} for a brief review) tries to find a reward function such that the associated optimal policy induces a distribution over trajectories (or some measure of this distribution) which matches the one induced by the expert: this is called apprenticeship learning. Most algorithms of apprenticeship learning use Reinforcement
Learning (RL) algorithms as subroutines, thus their performance depends strongly on the quality of these subroutines. Consequently, apprenticeship learning algorithms suffer from the
same challenges of large state spaces, exploration vs. exploitation trade-offs as RL. In order to avoid to solve repeatedly RL problems, we decide to take a different point of view.

Having in mind that there is a one to one relation between a reward function and its associated optimal quality function (via the inverse Bellman equation), it is worth thinking of a method which is able to output a quality function for which the greedy policy is the demonstrated expert policy. Thus, the demonstrated expert policy will be optimal for the reward function associated to the quality function. As a method which is able to output a quality function for which the greedy policy is the demonstrated expert policy, we propose a score-based classification (see Section~\ref{sec:algo}). Besides, in order to retrieve, via the inverse Bellman equation, the reward associated to the score function the classification outputs, we propose a regression (see Section~\ref{sec:algo}). That is why the method is called the Cascaded Supervised Approach (CSA). This method is justified by a theoretical analysis (see Section~\ref{sec:analysis}) where it is shown that the demonstrated expert policy is near-optimal for the reward the regression step outputs.

Finally, an instantiation of CSA is proposed~\ref{sec:experiments}. This algorithm does not need to solve iteratively an RL problem and needs only sampled transitions from expert and non-experts policies as inputs. Moreover, up to the use of some heuristics, the algorithm is able to be trained only with transitions sampled from the demonstrated expert policy. The algorithm is tested on a car driving simulator and compared to an apprenticeship learning algorithm, to a pure classification algorithm and to a recent successfull IRL method.

\section{Background and notations}
\label{sec:background}
%% \begin{itemize}
%% \item Parler du domaine de la prise de décision séquentielle
%% \item Expliquer le problème direct, en profiter pour introduire l'équation d'évaluation de Bellman et l'équation d'optimalité de Bellman
%% \begin{itemize}
%% \item $RL(MDP=\{S,A,\gamma,(P),R\}) = \pi^*:S\times A \rightarrow \mathbb{R}$
%% \item $\forall \pi, Q^\pi(s,a) = R(s,a) + \sum_{s'}P(s'|s,a)Q^\pi(s',\pi(s'))$
%% \item $\pi^* = \arg\max_aQ^*(s,a)$
%% \end{itemize}
%% \item Introduire le problème inverse dans une forme mathématique relativement propre
%% \begin{itemize}
%% \item $IRL( MDP\backslash R = \{S,A,\gamma,(P)\} \cup \{\pi^E\} = R : S\times A \rightarrow \mathbb{R}$ telle que $\pi^E$ est optimale pour $R$
%% \item Mentionner la nature mal posée du problème sans s'appesentir dessus. Préciser toutefois qu'aucune approche ne le résoud de manière mathématiquement satisfaisante à cause de ça.
%% \item insister sur l'optimalité de l'expert comme hypothèse. Un expert sous optimal, c'est un autre champ de littérature, éventuellement donner deux trois refs.
%% \end{itemize}
%% \end{itemize}
First, we introduce some general notations.
Let $E$ and $F$ be two non-empty sets, $E^F$ is the set of functions from $F$ to $E$.
We note $\Delta_X$ the set of distributions over $X$.
Let $\alpha\in\mathbb{R}^X$ and $\beta\in\mathbb{R}^X$: $\alpha\leq\beta \Leftrightarrow \forall x\in X, \alpha(x) \leq \beta(x)$. We will often slightly abuse the notations and consider (where applicable) most objects as if they were matrices and vectors indexed by the set they operate upon.

We frame the sequential decision problem in a finite \emph{Markov Decision Process} (MDP), that is a tuple $\{S,A,P,R,\gamma\}$. $S$ is a finite state space, $A$ is a finite action space, $R\in\mathbb{R}^{S\times A}$ is a reward-function, $\gamma\in ]0,1[$ is a discount factor and $P\in \Delta_{S}^{S\times A}$ is the dynamic of the MDP. Thus for each $(s,a)\in S\times A$, $P(.|s,a)$ is a distribution over $S$ and $P(s'|s,a)$ is the probability to reach $s'$ by choosing the action $a$ in the state $s$. At each time step $t$, the agent uses the information encoded in the state $s_t\in S$ in order to choose an action $a_t \in A$ according to a (deterministic\footnote{We restrict ourselves here to deterministic policies, but the loss of generality is minimal as there exists at least one deterministic optimal policy.}) \emph{policy} $\pi\in A^S$. The agent then transits to a new state $s_{t+1}\in S$ according to the markovian transition probabilities $P(s_{t+1}|s_t,a_t)$. The reward function $R$ is a local measure of the quality of the control. The global quality of the control induced by a policy $\pi$, with respect to a reward $R$ is assessed by the value function $V^\pi_R \in \mathbb{R}^{S}$ which associates each state with the discounted cumulative reward for following policy $\pi$ from this state: $V^\pi_R(s) = \E[\sum_{t\geq 0}\gamma^tR(s_t)|s_0 = s]$. Therefore, an optimal policy $\pi^*_R$ is a policy whose value function (the optimal value function $V^*_R$) is greater than that of any other policy, for all states : $\forall \pi, \forall s \in S, V^*_R(s) \geq V^\pi_R(s)$. The stationary distribution induced by a policy $\pi$ is written $\rho_\pi,\textrm{ s.t. }\rho_\pi^TP^\pi = \rho_\pi$.

The Bellman evaluation operator $T^\pi_R: \mathbb{R}^{S} \rightarrow  \mathbb{R}^{S}$ is defined by $T^{\pi_R}V = R + \gamma P_\pi V$ where $P_\pi$ is $(p(s'|s,\pi(s)))_{s,s' \in S}$. The Bellman optimality operator follows naturally : $T^*_RV = \max_\pi T^\pi_RV$. Both operators are contractions. The fixed point of the Bellman evaluation operator $T^\pi_R$ is the value function of $\pi$ with respect to reward R: $V^\pi_R$; the fixed point of the Bellman optimality operator $T^*_R$ is the optimal value function $V_R^*$ with respect to reward $R$. Optimal value function and optimal policy are also linked via the quality function $Q^\pi_R\in\mathbb{R}^{S\times A}$, a state-action value function that adds a degree of freedom on the choice of the first action, formally defined by $Q^\pi_R(s,a) = T^a_RV^\pi_R(s)$, with $a$ the policy that always returns action $a$ ($T^aV = R + \gamma P_a V$ with $P_a = (p(s'|s,a))_{s,s' \in S}$). The optimal policy follows a greedy mechanism with respect to its optimal quality function 
\begin{equation}
  \label{eq:greedy}
  \pi^*_R(s)\in\argmax_aQ^*_R(s,a).
\end{equation}

When the state space is too large to allow matricial representation or when the transition probability or even the reward function are unknown except through observations gained by interacting with the system, RL or approximate dynamic programming may be used to retrieve the optimal control policy \cite{sutton1998reinforcement}. We call this the direct problem.

In this work, we wish to tackle the inverse problem. We observe an expert's deterministic\footnotemark[\value{footnote}] policy $\pi^E$, assuming that there exist some unknown reward $R^E$ for which the expert is optimal. The non optimality of the expert is an interesting setting that has been discussed for example in \cite{meloXXXanalysis} but that we are not addressing here. We do not try to find this unknown reward $R^E$ but rather a non trivial reward $R$ for which the expert is also near-optimal. The trivial reward $0$ is a solution to this ill-posed problem (no reward means that every behavior is optimal). Because of its ill-posed nature, this expression of \emph{Inverse Reinforcement Learning} (IRL) still has to find a satisfactory solution although lots of progress have been made, see section \ref{sec:related}.
\section{The cascading algorithm}
\label{sec:algo}
%% \begin{itemize}
%% \item Mentionner (sans en parler, ce sera fait dans \hyperref[sec-7-1]{Imitation}) l'imitation. Une approche supervisée évidente est la classification, avec pour entrée un set $D_C$ de données issues de l'expert. Rappeler (expliqué pour la première fois dans \hyperref[sec-3]{Formalisme} qu'on se restreint (sans perte de généralité, le préciser) au cas des politique déterministes.
%% \item Tous les classifieurs utilisent une fonction de score (je n'ai pas trouvé un seul exemple d'un classifieur qui n'en utilise pas), il s'agit juste d'un manière d'écrire les choses.
%% \item Expliquer que l'on identifie le mécanisme de classification au mécanisme glouton qui régit la politique optimale (donc celle de l'expert).
%% \item Sans chercher à justifier mathématiquement la chose (ce sera fait dans la partie \hyperref[sec-5]{Analyse}), expliquer qu'intuitivement l'on va identifier fonction de score du classifieur et fonction de qualité de l'expert (qui est optimale pour une certaine récompense, rappelons-le).
%% \begin{itemize}
%% \item Repartir de l'équation d'évaluation de Bellman pour isoler le terme $R(s,a)$ et introduire $r:S\times A\in \mathbb{R}$, la récompense issue de cette équation lorsque l'on remplace $Q^E$ par la fonction de score : $r(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi(s')) $
%% \item On estime ce terme à partir des transitions $D_R$ car les probabilités de transitions sont inconnues : $\hat r(s,a) = q(s,a)-\gamma \sum_{s'} {1\over N_{s'}}\sum_{(s,a,s',a')\in D_R}q(s',a')$
%% \item La fonction de récompense est donnée par une régression sur $D_R \cup {\hat r}$.
%% \end{itemize}
%% \item Donner l'algorithme l'environnement \texttt{algorithm2e}
%% \item Rapidement, l'étape de régression permet d'ajouter la structure du MDP au résultat de classif, en utilisant l'équation de Bellman.
%% \item Bien préciser qu'il ne s'agit ici que d'une explication qualitative, ``avec les mains'', la preuve du fait que ça résoud bien le problème est donnée dans \hyperref[sec-5]{Analyse}, une démonstration sur des problèmes si complexes que l'état de l'art ne peut les résoudre (sauf SCIRL) est donnée dans \hyperref[Experience]{Experience}. Les avantages comparés sont expliqués dans \hyperref[Autres-Approches]{Autres Approches}. Enfin, des explications plus intuitives sur les mécanismes internes de l'algorithme, ainsi que des résultats expérimentaux supplémentaires sont fournis en \hyperref[sec-9]{Annexes}.
%% \end{itemize}
Our first step in the quest for a reward function solving the IRL problem is a classification step using a \emph{score function-based multi-class classifier} (MC$^2$). This classifier learn a score function $q:S\times A \rightarrow \mathbb{R}$ that rates the association of a certain label $a\in A$ with a certain input $s\in S$. The classification rule $\pi_C: S \rightarrow A$ is simply to select (one of) the label(s) that achieve the highest score for the given input :
  \label{eq:greedy2}
\pi_C(s) \in \argmax_a q(s,a).
Most notably, \emph{Multi-class Support Vector Machines} can be seen as MC$^2$ algorithms. We give more examples and detail our choice of a MC$^2$ algorithms for the experiments in subsection \ref{sec:instantiation}.

Given a dataset $D_C = \{(s_i,a_i=\pi^E(s_i))_i\}$ of actions $a_i$ (deterministically) chosen by the expert on states $s_i$, we train such a classifier. Apart from discussion and comparison purpose, we do not make use of the classification policy $\pi_C$. What is of interest to us is the score function $q$ itself. One can easily notice the similarity between equations \ref{eq:greedy2} and \ref{eq:greedy} that describes the relation between the optimal policy in a MDP and its optimal quality function. As the expert is optimal with respect to some unknown reward function $R_E$ and we seek a reward function $R$ for which it is also optimal, it is only natural to view the score function $q$ of the classifier as some kind of optimal quality function for the classifier policy $\pi_C$. By inversing the Bellman evaluation equation, one can express the reward as a function of the quality function : $R(s,a) = Q^*_R(s,a) - \sum_{s'}P(s'|s,a)Q^*_R(s'\pi^*(s'))$. This leave us with $r$, the reward function relative to our score/quality function $q$ : $r(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi_C(s'))$.

We wish to solve the general, or approximate, IRL problem where the transition probabilities are unknown. For this we use another dataset $D_R = \{(s_j,a_j,s'_j)_j\}$ where $s'_j$ is the state an agent taking action $a_j$ in state $s_j$ transitioned to. Action $a_j$ need not be chosen by any particular policy. The dataset $D_R$ brings us information about the dynamics of the system. Our reward function $r$ must be approximated to $\hat r$ with the help of the information in $D_R$. This is done via datapoints $\{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j$ that we then feed to a regressor FIXME:citation 36/15 papier de munos to obtain $R^C$, a generalisation of $\hat r$ over the whole state-action space. The complete algorithm is given Alg. \ref{algo:cascading}.
\begin{algorithm}%[H]
    %\small
  \caption{Cascading IRL algorithm}
  \label{algo:cascading}
  \emph{\textbf{Given}} a training set $D_C=\{(s_i,a_i=\pi^E(s_i))\}_{1\leq i \leq D}$ and another training set $D_R=\{(s_{j},a_{j},s'_{j})\}_{1\leq j \leq D'}$\;\\
  \emph{\textbf{Train}} a score function-based classifier on $D_C$, obtaining decision rule $\pi_C$ and score function $q:S\times A \rightarrow \mathbb R$\;\\
  \emph{\textbf{Learn}} a reward function $\hat R^C$ from the dataset $\{(s_{j},a_{j},\hat{r}_j)\}_{1\leq j \leq D'}$, $\forall (s_j,a_j,s'_j) \in D_R,\hat{r}_j=q(s_{j},a_{j})-\gamma q(s'_{j},\pi_C(s'_{j}))$\;\\
  \emph{\textbf{Output}} the reward function $\hat R^{C}$ \;
\end{algorithm}

Cascading two supervised approaches like we do is a way to inject the MDP structure in the resolution of the problem. Indeed, mere classification only takes into account information from the expert (i.e., which actions goes with what state), but there is more to a MDP than a single policy. Using the Bellman equation in the expression of $\hat r$, we make use of the information lying in the transitions $(s,a,s')$, namely information about the transition probability $P$. The final regression step is just a way to try to generalize this information about $P$ to the whole state-action space, in order to have a well-behaved reward function.

As the analysis (Sec \ref{sec:analysis}) reveals, both classification error and regression error should be controlled. By carefully choosing or tuning both supervised algorithm, it should possible to mitigate any ill effects of the specific settings the cascading algorithm is used in. If one operates in a well-known domain where a lo of data is available, then ease-of-use will probably dictate the choice of two off-the-shelf implementations with default settings. If data is lacking, one could for example have to trade off between bias and variance. The possibility to use and leverage the wide range of techniques developped for these two well known supervised problems is a strong advantage of the proposed cascading approach.
%Reformuler ce passage
The explanation in this section only is a quatitative attempt to make the reader understand the idea behind the algorithm. The analysis of the next section tackle things in a formal, yet less intuitive, way. In Sec \ref{sec:experiments}, the use of heuristics (not tackled in the analysis) is empirically shown to be relevant. Data constraints are so high that to the best of our knowledge only one other published algorithm is able to solve the IRL problem in such conditions. Other approaches are compared with one another in Sec. \ref{sec:related}. Finally, appendices provide the reader with other experiments that exemplify the inner working of the algorithm.

@@

In this Section, we prove that the deterministic expert policy $\pi_E$ is near optimal for the reward $\hat{R}^C$ outputted by the regression step. We formulate this result in a more formal way in Theorem \ref{theorem: theorem analysis}

\begin{theorem}
\label{theorem: theorem analysis}
Let $\pi_E$ be the deterministic expert policy, $\rho_E$ its stationary distribution and $\hat{R}^C$ the reward outputted by the regression step. We have:
$\Delta q$, and $C_*$ are constants and are defined in the proof. $\epsilon_C$ represents the classification error and $\epsilon_R$ represents the regression error: both terms are well-defined in the proof.
\end{theorem}
\begin{proof}
First let introduce some notations. $q\in\mathbb{R}^{S\times A}$ is the score function outputted by the classification step, $\forall s \in S,\pi_C\in\argmax{a\in A}q(s,a)$ is a deterministic classifier policy and $\epsilon_C$ is the classification error:

@@

Moreover we recall the definition of the reward $R^C\in\mathbb{R}^{S\times A}$:
\forall (s,a)\in S\times A, R^C(s,a)=q(s,a) +\gamma\sum_{s'\in S}P(s'|s,a)q(\pi_C(s'),a),
and the definition of $\hat{R}^C\in\mathbb{R}^{S\times A}$ which is the reward function outputted by the regression step.
The difference between $R^C$ and $\hat{R}^C$ is noted $\epsilon^R=R^C-\hat{R}^C$, $\epsilon^R\in\mathbb{R}^{S\times A}$ must not be mistaken for the regression error $\epsilon_R\in\mathbb{R}$ defined below.
We also introduce the reward function $R^E\in\mathbb{R}^{S\times A}$ which will be useful in our proof:
\forall (s,a)\in S\times A, R^E(s,a)=q(s,a) +\gamma\sum_{s'\in S}P(s'|s,a)q(\pi_E(s'),a).

@@

&R^E_a=q_a+\gamma P_aq_{\pi_E},

@@

Now, there is a still an upper bound to find for the second term. It is possible to decompose the second term:
